\documentclass[12pt]{report}

% packages used for many things
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan}
\urlstyle{same}
% package used for the enumeration
\usepackage{enumitem}
% packages used to write more symbols and text in math mode
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsfonts} 
\usepackage{amssymb}
\usepackage{MnSymbol}
\usepackage{csquotes}
\usepackage{arydshln}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{extarrows}
\usepackage{tikz}


\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=15mm,
 }


% for images
\usepackage{graphicx}
\graphicspath{{./}}

% qed black square
\newcommand*{\QEDA}{\hfill\ensuremath{\blacksquare}}
% xor symbol
\newcommand*\xor{\oplus}

\title{Artificial Intelligence II \\ Assignment 5 Report}
\author{Andreas - Theologos Spanopoulos (sdi1700146@di.uoa.gr)}
\date{December 31, 2020}


% ----------------    START OF DOCUMENT    ------------ %
\begin{document}
\maketitle

\section*{Questions 1-25}
\begin{enumerate}
    \bigskip
    % ---------------------------       1-10       --------------------------- %
    \item How many McDonald's stores are there in the capital of Ireland?
    \item From the cities bordering Liverpool, which one has the biggest area?
    \item Which airport is closer to the centre or Chester?
    \item How many students are registered in the University College of Dublin?
    \item Is there a river near the Woodlesford Primary school?
    \item Which churches reside in 1km radius from Margaret McMillan Park?
    \item Are there Gift Shops inside the Flamingo Land zoo?
    \item Which coffee shops are by the Union Canal in Edinburgh?
    \item Where is the closest ATM to the Hotel Le Monde, Edinburgh?
    \item In which town is the Highland Hospice Hospital located?
    % ---------------------------       11-20       --------------------------- %
    \item Which Hospital is located between the Summefield Park and the Birmingham Canal?
    \item Is there Parking in the Cardiff University?
    \item Is the Sherman Theatre closer the the Cardiff University than the National Museum of Cardiff?
    \item Which mountains of Newport are higher than the highest mountain of Bristol?
    \item Are there Traffic Signals next to the Harpal Clinic in London?
    \item Where is the closest Bus Station from the Shoe Lane Library in London?
    \item What is the area of the Hyde Park in London?
    \item Does the London Zoo have a Bus Station next to it?
    \item Which is the longest Cycleway in Dublin?
    \item Is there a Zoo in Dublin?
    % ---------------------------       21-25       --------------------------- %
    \item Is there a Post Office close to the Cork City Library?
    \item Does the South Mall in Cork city contain a Pharmacy and a Grocery Store?
    \item How many Golf Fields are there in Cork city?
    \item How far is the Applegreen Gas Station from the St. Michaels church in Tipperary?
    \item Are there Italian Restaurants in 3km radius from the Cahir castle in Ireland?
\end{enumerate}
    \clearpage

\section*{Questions 26-50}
\begin{enumerate}
    \setcounter{enumi}{25}
    \bigskip
    % ---------------------------       26-30       --------------------------- %
    \item Which are the Nightclubs in Manchester that have a Hostel in 500m radius?
    \item Which is the name of the Football team that has its stadium in the city
        which the Oxford University belongs to?
    \item Is there a Public Toilet next to the Cineworld cinema in Nottingham?
    \item Where to buy sports equipment in Leeds?
    \item Is there a fountain in the Hoglands Park in Southampton?
    % ---------------------------       31-40       --------------------------- %
    \item What is the length of the Delaware River?
    \item Does the Abraham Lincoln High School have a swimming pool?
    \item What is the height difference of the Statue of Liberty and the Empire
        State Building?
    \item Which Boroughs of New York City does the 'New York City Subway' connect?
    \item Which brands have Mobile Phone Shops both in Manhattan, Queens and Princeton?
    \item Which Parks in Orlando have small lakes inside?
    \item What are the opening hours of the Kmart building in Key West?
    \item How many beaches are there in the coast of Miami?
    \item Which cities of Los Angeles have a population bigger than 10000 people?
    \item What is the name of the NBA team that resides in the Los Angeles area?
    % ---------------------------       41-50       --------------------------- %
    \item Which Forests of Honolulu have a protection class equal to 4?
    \item Which Dog Parks have a small Lake nearby in Chicago?
    \item Which University in Illinois has the most post-doctoral students?
    \item Is the Berkeley University located in California?
    \item What is the height difference between the 2 highest peaks of San Francisco?
    \item Are there Greek Restaurants between the 2 airports of San Jose?
    \item In which district of Washington does the University of Washington belong to?
    \item Does Lake Sammarish cover a smaller area than Phantom Lake?
    \item Does the Forbes Avenue go through the Carnegie Mellon University?
    \item Which companies have offices in Silicon Valley?
\end{enumerate}

\end{document}
% ----------------    END OF DOCUMENT    ------------ %
